---
title: "CMCC_assignment_01"
output: html_notebook
---

# BONEAU REPLICATION
On the robustness of the T-Test.

- N = (5, 15)
- VARIANCE = (1, 4)
- SAMPLING OF CELLS = (Uniform, Normal, Exponential)

Note that the `mean = 0` for all distributions

## Function: Compute T Values
```{r}

compute_t_vals <- function(dist1, dist2, iterations){
  
  Ns = list(c(5,5), c(5,15), c(15,15))
  variances = list(c(1,1), c(1,4), c(4,4))
  
  # initialize matrix
  t_vals = matrix(ncol = iterations, nrow = 9)
  # initialize a dataframe to hold parameters
  params= data.frame(dist1 = NA, dist1_N = NA, dist1_var = NA, dist2 = NA, dist2_N = NA, dist2_var = NA,
                     theoretical_95 = NA, actual_95 = NA)
  # create a counter
  counter = 1
  
  for (N in Ns){
    for (variance in variances){
  
      # calculate t values for parameters
      t_vals[counter,] = t_calcs(N, variance, dist1, dist2)
      
      # create list based on parameters
      t_95 = calc_middle(N)
      a_95 = sum(abs(t_vals[counter,]) > t_95)/iterations
      params = rbind(params, c(dist1, N[1], variance[1], dist2, N[2], variance[2], t_95, a_95))
      
      # add 1 to counter
      counter=counter+1
    }
  }
  # clean up params
  params = params[-1,]
  rownames(params) <- seq(length=nrow(params))
}

```

## Function: Calculate sample % outside 95% t-value
```{r}
calc_middle <- function(N){
  # 8 DOF
  if (N[1]+N[2] == 10){
    val = 2.262
  }

  # 18 DOF
  if (N[1]+N[2] == 20){
    val = 2.101
  }

  # 28 DOF
  if (N[1]+N[2] == 30){
    val = 2.048
  }
  return(val)
}


```

## Function: Calculate distributions
```{r}
# FUNCTION TO CALCULATE T VALS
t_calcs <- function(N, variance, dist1, dist2){
  
  # check dist for first sample
  if (dist1 == 'norm'){
    sample1 = replicate(iterations, rnorm(N[1], mean=0, sd=sqrt(variance[1])))
  }
  else if(dist1 == 'exp'){
    sample1 = replicate(iterations, rexp(N[1], mean=0, sd=sqrt(variance[1])))
  }
  else if(dist1 == 'unif'){
    sample1 = replicate(iterations, runif(N[1], mean=0, sd=sqrt(variance[1])))
  }
  
  # check dist for second sample
  if (dist1 == 'norm'){
    sample2 = replicate(iterations, rnorm(N[2], mean=0, sd=sqrt(variance[2])))
  }
  else if(dist1 == 'exp'){
    sample2 = replicate(iterations, rexp(N[2], rate = )-1)
  }
  else if(dist1 == 'unif'){
    sample2 = replicate(iterations, runif(N[2], mean=0, sd=sqrt(variance[2])))
  }
  
  t_vector = rep(NA, iterations)

  for (i in 1:iterations){
    t_vector[i]= t.test(sample1[,i],sample2[,i], var.equal = TRUE)$statistic 
  }
  return(t_vector)
}
```

## Function: Plot a Frequency Distribution of the Obtained Ts

```{r}
# which row

plot_tests <- function(param_combo, t_vals, params){

  dof = as.numeric(params$dist1_N[param_combo]) + as.numeric(params$dist2_N[param_combo])
  x <- seq(-4, 4, length=100)
  
  hist(x=t_vals[param_combo,],
       breaks=50,
       freq = FALSE,
       xlab = "Value of t",
       main = paste0("DIST 1: ", params[param_combo,1], ", n=", params[param_combo,2], ", v=", params[param_combo,3],
                     "  ||  DIST 2: ", params[param_combo,4], ", n=", params[param_combo,5], ", v=", params[param_combo,6]),
       sub = paste0("% beyond 95% center: ", params$actual_95[param_combo])
       )
  lines(x, dt(x, dof-2), lwd=2)
  abline(v = as.numeric(params$theoretical_95[param_combo]), col="red")
  abline(v = -1 * as.numeric(params$theoretical_95[param_combo]), col="red")
}


```

## PLOT A VERSION OF A CELL

```{r}

compute_t_vals('norm', 'norm', 1000)

# PLOT
par(mfrow=c(3,3))

for (i in 1:9){
  plot_tests(i, t_vals, params)
}

```

Could be interesting to look at distribution of t-test error % - if you run it a bunch should get close to 5%